### 2018-03-23 Sections/quantum_genetic_medicine.tex
tiny
Gen med figs

To achieve this, there are two approaches that are commonly employed -- \textit{de novo} assembly, and \textit{mapping}.

@@

The number of comparisons between short-reads scales only as $O(N^2)$, and employing hash-table\index{Hash-table} representations for short-reads, the lookups for each individual comparison can be efficiently implemented in $O(1)$ time. The confidence that two contigs/reads actually overlap increases exponentially with $n$, but with increasing $n$ comes a reduction in the number of matches that will be identified -- the tradeoff between \textit{sensitivity}\index{Sensitivity} and \textit{specificity}\index{Specificity} in contig reconstruction.

@@

@@

@@

These Hamiltonians can be constructed \textit{ab initio}, combining relevant one- and two-body interactions\index{One-body interactions}\index{Two-body interactions}, such as Coulomb potential\index{Coulomb potential}, kinetic energy\index{Kinetic energy}, spin- and spin-orbit\index{Spin coupling}\index{Spin-orbit coupling} coupling, and magnetic and electric dipole terms appropriately. A net interaction Hamiltonian may be obtained by summing over the different one- ($i$) and two-body (\mbox{$j,k$}) terms for all particles comprising the system,
\begin{align}
	\hat{H}_\mathrm{int} = \sum_i \hat{H}_i + \sum_{j>k} \hat{H}_{j,k}.
\end{align}
This structure is of the form of Eq.~(\ref{eq:Ham_sim_Ham}) and thus lends itself to the quantum Hamiltonian simulation algorithm\index{Hamiltonian simulation} (Sec.~\ref{sec:quantum_sim_alg}).

The dominant one-body interactions are kinetic energy\index{Kinetic energy Hamiltonian} terms, of the form,
\begin{align}
\hat{H}_i^{(\mathrm{kinetic})} &= \frac{\hat{p}^2}{2m} \nonumber \\
&= -\frac{\hbar^2}{2m_i}\nabla^2_{\vec{r}_i},
\end{align}
where $\hat{p}$ is the momentum operator, $m_i$ is mass, $\vec{r}_i$ is the position vector, and,
\begin{align}
\nabla^2 = \frac{\partial^2}{\partial x^2} + \frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2}.
\end{align}
The dominant two-body terms are Coulomb (electrostatic) interactions\index{Coulomb interaction},
\begin{align}
\hat{H}_{j,k}^{(\mathrm{Coulomb})} = \frac{q_j q_k}{4\pi\epsilon_0|\vec{r}_j-\vec{r}_k|},
\end{align}
where $q_j$ denotes charge.

@@

To achieve this, let us construct a classical algorithm for exhaustively constructing and enumerating organic drug molecules or functional groups, up to some cutoff size\index{Cutoff size}. The length of this list grows exponentially with the cutoff size. Next we implement this algorithm unitarily (which is always possible, writing the classical circuit as a reversible one), and construct it as a quantum oracle\index{Oracle} such that every input state (which acts as a pointer\index{Pointers} to an element in the list) yields as output a qubit representation\index{Qubit molecule representation} of the corresponding drug compound. Specifically, the oracle implements a transformation of the form,
\begin{align}
\hat{U}_\mathrm{oracle}\ket{i}_\mathrm{pointer}\ket{0}_\mathrm{drug}^{\otimes n} \to \ket{i}_\mathrm{pointer}\ket{\psi_i}_\mathrm{drug}^{(n)},
\end{align}
where $i$ is the index (pointer) to the drug element in the enumeration, and $n$ is the number of qubits encoding the representation of the drug molecule\footnote{This may be implemented unitarily, which is easy to see via its preservation of orthonormality.}.

@@

Having constructed this oracle, we feed the output qubit molecule description into the Hamiltonian simulation algorithm\index{Quantum simulation}, with exponential speedup, yielding as output a `score' characterising some desired property of the interaction, such as bonding affinity\index{Bonding affinity}. This implements the simulate-and-score unitary,
\begin{align}
\hat{U}_\mathrm{sim} \ket{\psi_i}_\mathrm{drug}\ket{0}_\mathrm{score} \to \ket{\psi_i}_\mathrm{drug}\ket{s_i}_\mathrm{score}	
\end{align}

>>>>

To achieve this, there are two approaches that are most commonly employed -- \textit{de novo} assembly, and \textit{mapping}.

@@

The number of comparisons between short-reads scales only as $O(N^2)$, and employing hash table\index{Hash table} representations for short-reads, the lookups for each individual comparison can be efficiently implemented in $O(1)$ time. The confidence that two contigs/reads actually overlap increases exponentially with $n$, but with increasing $n$ comes a reduction in the number of matches that will be identified -- the tradeoff between \textit{sensitivity}\index{Sensitivity} and \textit{specificity}\index{Specificity} in contig reconstruction.

@@

Fig.~\ref{fig:gen_seq_de_novo} provides a graphic example of \textit{de novo} sequencing.

\begin{figure}[!htb]
	\includegraphics[width=0.47\textwidth]{genetic_sequencing_de_novo}
	\caption{Three short-reads, each of 10 base-pairs in length (blue), which we wish to assemble into the largest possible contig belonging to the underlying genome. The specificity\index{Specificity} defines the number of endpoint base-pairs that must overlap for them to be considered a match (orange). In this case the specificity is 3, and 3 short-reads have been assembled into a single 24 base-pair contig. The greater the specificity the more accurate will be the matching, but the lower the probability of finding matches. A specificity too low will create false matches, whereas a specificity too high makes finding sufficiently overlapping short-reads unlikely. This makes optimum choice of specificity an important consideration, creating a direct tradeoff between reconstructed contig length and accuracy. This sequencing technique is best applied when there is no known reference genome for the data under analysis. For the technique to work effectively there must be a sufficient quantity of short-read data that with high probability overlaps between reads will exist. If the data is too sparse no overlaps are likely to be found.} \label{fig:gen_seq_de_novo}
\end{figure}

@@

Fig.~\ref{fig:gen_seq_mapping} provides a graphic example of mapping in the presence of a SNP.

\begin{figure}[!htb]
	\includegraphics[width=0.47\textwidth]{genetic_sequencing_mapping}
	\caption{In genome mapping we begin with a known reference genome (blue) to act as a comparison for short-reads (orange). The short-reads are slided along the reference until a location with sufficient overlap is found, at which point we record a match. By allowing some degree of single nucleotide mismatches (green) between the reference and short-read, we are able to characterise the SNPs in the short-read data. This sequencing technique is appropriate when we have a known genome for the species to which the short-read data belongs, but we wish to identify minor genetic mutations between the samples. For example, mapping short-reads belonging to human cancer cells to a human reference genome characterises the genetic mutations associated with that particular cancer. The technique works with any amount of short-read data, even just a single read, provided the reference is known.} \label{fig:gen_seq_mapping}
\end{figure}

@@

These Hamiltonians can be constructed \textit{ab initio}, combining relevant one- and two-body interactions\index{One-body interactions}\index{Two-body interactions}, such as Coulomb potential\index{Coulomb potential}, kinetic energy\index{Kinetic energy}, spin- and spin-orbit\index{Spin coupling}\index{Spin-orbit coupling} coupling, and magnetic and electric dipole terms appropriately. A net interaction Hamiltonian may be obtained by summing over the different one- ($i$) and two-body (\mbox{$j,k$}) terms for all particles comprising the system,
\begin{align}
	\hat{H}_\text{int} = \sum_i \hat{H}_i + \sum_{j>k} \hat{H}_{j,k}.
\end{align}
This structure is of the form of Eq.~(\ref{eq:Ham_sim_Ham}) and thus lends itself to the quantum Hamiltonian simulation algorithm\index{Hamiltonian simulation} (Sec.~\ref{sec:quantum_sim_alg}).

The dominant one-body interactions are kinetic energy\index{Kinetic energy Hamiltonian} terms, of the form,
\begin{align}
\hat{H}_i^{(\text{kinetic})} &= \frac{\hat{p}^2}{2m} \nonumber \\
&= -\frac{\hbar^2}{2m_i}\nabla^2_{\vec{r}_i},
\end{align}
where $\hat{p}$ is the momentum operator, $m_i$ is mass, $\vec{r}_i$ is the position vector, and,
\begin{align}
\nabla^2 = \frac{\partial^2}{\partial x^2} + \frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2}.
\end{align}
The dominant two-body terms are Coulomb (electrostatic) interactions\index{Coulomb interaction},
\begin{align}
\hat{H}_{j,k}^{(\text{Coulomb})} = \frac{q_j q_k}{4\pi\epsilon_0|\vec{r}_j-\vec{r}_k|},
\end{align}
where $q_j$ denotes charge.

@@

To achieve this, let us construct a classical algorithm for exhaustively constructing and enumerating organic drug molecules or functional groups, up to some cutoff size\index{Cutoff size}. The length of this list grows exponentially with the cutoff size. Next we implement this algorithm unitarily (which is always possible, writing the classical circuit as a reversible one), and construct it as a quantum oracle\index{Oracle} such that every input state (which acts as a pointer\index{Pointers} to an element in the list) yields as output a qubit representation\index{Qubit molecule representation} of the corresponding drug compound. Specifically, the oracle implements a transformation of the form,
\begin{align}
\hat{U}_\text{oracle}\ket{i}_\text{pointer}\ket{0}_\text{drug}^{\otimes n} \to \ket{i}_\text{pointer}\ket{\psi_i}_\text{drug}^{(n)},
\end{align}
where $i$ is the index (pointer) to the drug element in the enumeration, and $n$ is the number of qubits encoding the representation of the drug molecule\footnote{This may be implemented unitarily, which is easy to see via its preservation of orthonormality.}.

@@

Having constructed this oracle, we feed the output qubit molecule description into the Hamiltonian simulation algorithm\index{Quantum simulation}, with exponential speedup, yielding as output a `score' characterising some desired property of the interaction, such as bonding affinity\index{Bonding affinity}. This implements the simulate-and-score unitary,
\begin{align}
\hat{U}_\text{sim} \ket{\psi_i}_\text{drug}\ket{0}_\text{score} \to \ket{\psi_i}_\text{drug}\ket{s_i}_\text{score}	
\end{align}
